---
title: "Breast Cancer Analysis"
author: "Me"
date: "5/25/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Exploration

The following chunk downloads and reads the data, then changes the names to sensible ones. The top 6 rows are then printed. 

```{r message=FALSE}
library(dplyr); library(ggplot2)
url <- "https://archive.ics.uci.edu/ml/machine-learning-databases/breast-cancer-wisconsin/breast-cancer-wisconsin.data"
download.file(url = url, destfile = "./breastcancerdata.csv")
cancer.data <- read.csv("./breastcancerdata.csv")
names(cancer.data) <- c("Sample.code","Clump.Thickness", "Cell.Size.Unif","Cell.Shape.Unif",
                        "Marginal.Adhesion","Epit.Cell.Size","Bare.Nuclei","BlandChrom",
                        "NormalNucleoli","Mitoses", "Class")
head(cancer.data)

```

I also added a column to clarify malignant v. benign and got rid of a couple of columns, then made the added column into a factor variable. 

```{r}

cancer.data <- cancer.data %>% 
  mutate(Class.Name = ifelse( Class == 2, "Benign", "Malignant")) %>%
  select(2:10,12)
cancer.data$Class.Name <- factor(cancer.data$Class.Name)

```

Now then, let's look at some summary statistics

```{r}

summary(cancer.data)

```

## Short Analysis

I'm going to go for low-hanging fruit and look at just one of the predictors, clump thickness. So I'll squish down my columns even more:

```{r}

cancer.data <- select(cancer.data, Clump.Thickness, Class.Name)

```

What sort of effect may be going on? Let's make a boxplot to find out:

```{r}

g <- ggplot(cancer.data, aes(Class.Name, Clump.Thickness))
g + geom_boxplot()

```

Looks like a definite difference; let's make histogram to illustrate it differently:

```{r}

g <- ggplot(cancer.data, aes(Clump.Thickness))
g + geom_histogram(aes(fill = Class.Name), bins = 10) 
            

```

Just for fun, I'm also going to build a quick decision tree:

```{r message = FALSE}

library(rpart)

fit <- rpart(Class.Name ~ Clump.Thickness, method = "class", data = cancer.data)
summary(fit)

```

As shown by my little tree, I have a split point at Clump Thickness of 6.5; anything below is predicted as Benign, anything above is predicted as Malignant. It's not a perfect model, but it's a useful coarse predictor. And it shows that I expect that with increasing clump thickness, there is incresing likelihood of malignancy, with over 96% probability of malignancy over a classification of 6.5. 