---
title: "Week 3 Solution"
author: "Matt Richards"
date: "7/24/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Here's one possible solution to the in-class exercises:

We'll start by reading in our files:

```{r message = FALSE, warning = FALSE}
library(dplyr)
human.hoco <- read.table(file = "../Datasets/tomtom_human.txt")
mouse.hoco <- read.table(file = "../Datasets/tomtom_mouse.txt")
all.hoco <- rbind(human.hoco,mouse.hoco)
names(all.hoco) <- c("Query.ID", "Target.ID", "Optimal.offset", "Pvalue","Evalue","Qvalue","Overlap","Query.consensus","Target.consensus","Orientation")
dim(all.hoco)
```

As we can see, there's about 30,000 records here. I'm now going to run through a series of commands that:

* Grab just the relevant columns
* Add a logP value
* Filter out only the logP values above my threshold of 7.3

I'm going to do this using pipes (`%>%`) which are introduced later in the `dplyr` lessons. If this is confusing now, I recommend reading ahead in DataCamp.

```{r}
above.p <- all.hoco %>% 
  select(Query.ID, Target.ID, Pvalue) %>%
  mutate(logP = -1*log10(Pvalue)) %>%
filter(logP >= 7.3)
head(above.p)
```

As we can see, I now have a mapping of queries from HOCOMOCO to targets from Jaspar. I 