---
title: "Week 2 Materials"
author: "Matt Richards"
date: "7/13/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Week Overview

Week 2, the second and final week of the DataCamp "Intermediate R" course, was sort of a mishmash of things. After starting with some very general concepts, we had sort of a miscellaneous section to end the section. Here are the big things we covered:

* Functions: one of the most central concepts in all programming. Functions let you build modular pieces of code that you can then apply to lots of different scenarios. 

* `lapply`: the commonly used "loop" function; you use it to apply a function to a list or vector and you **ALWAYS** get a list in return. 

* `sapply/vapply`: variations on lapply that return other data types. The `sapply` simplifies it however it sees fit, so it can be a little less predictable than `vapply`, where you specify exactly the data type you want. 

* `abs/sum/mean/round`: various mathematical utilities that do precisely what they say. These will come in handy all over the place as you build functions and scripts

* `seq/rep/sort/rev/str/append/is.*/as.*/unlist`: various data utilities that allow you to generate, manipulate, and evaluate your data. Like the math utilities, these will frequently pop up in your functions and scripts

* Regular expressions: a language of literal characters + metacharacters that allow you to search for and replace complex patterns in your data. 

* `grep/grepl`: functions that match an expression to your data. The `grep` returns indices or values, whereas the `grepl` returns a logical vector

* `sub/gsub`: functions that find and substitute an expression in your data. The `sub` does so for just the first occurence whereas the `gsub` does it for all occurences

* Times/Dates: R has built-in data types to deal with times and dates, allowing you to figure out elapsed time and things of that nature. You can convert to/from dates, but you have to be careful of formatting. 

## Additional Material: the `apply` functions

I don't have much to add on functions; the oddest thing about the functions is getting used to default arguments and how they work. Instead, I'll briefly add some material about the `apply` family and then move on to a more involved discussion of regular expressions. 

In addition to the 3 `apply` type functions you learned in DataCamp, there's 2 more that might come in useful. One is literally `apply`; it's a useful command if you're dealing with a matrix of data as it lets you perform a function on either the columns or rows of the matrix. For instance, say I wanted to take the product of each column; I could use `apply` to do that:

```{r}
M <- matrix(1:10,nrow=2)
M
apply(M, 2, prod)
```

The second argument here specifies the "margin"; that's 1 for rows, 2 for columns. So what I've said here is take the product (`prod`) of each column of M. 

The other `apply` family member that might interest you is `tapply`, which comes from "table apply". We use `tapply` when we have data in separate groups where we want to perform an operation, but separate the results by group. Confused? Here's an example: suppose we have a data frame of patients who are either "Healthy", or "Infected", plus their ages. Something like this:

```{r}
set.seed(11)
patients <- data.frame(state = sample(c("Healthy","Infected"),20, replace = TRUE),
                       age = sample(30:50, 20, replace = TRUE))
patients
```

Now let's suppose I want the mean age of each group so that I can compare them. Using `tapply` does this very easily:

```{r}
tapply(patients$age, patients$state, mean)
```

Finally, also along the lines of `apply` is the `split` function, which splits a dataset into groups. If you use `split` on your dataset, you can then use other `apply` functions and get `tapply`-like results:

```{r}
split.patients <- split(patients$age, patients$state)
lapply(split.patients, mean)
```

## In-Class Exercises (Part 1)

For the first part of today's exercises, we'll specifically use functions and the `apply` family, though you're welcome to bring in other tools if you would like. 

1) Load the built-in R data set "airquality" as follows:

```{r}
data(airquality)
```

These data measure various things having to do with the air quality in New York. For more information, use the `?` operator:

```{r eval=FALSE}
?airquality
```

2) The data include a column called "Month" where each month is encoded as an integer. Using the skills you've learned in Intermediate R, add a new column to the data frame called "Month.Name" that contains the actual name of the month (e.g. 1 becomes January). 

3) Now that you have month names in your data frame, find the mean of the "Ozone" variable for each month. Which month has the highest mean Ozone level? Which is lowest?

## Additional Material: Regular Expressions and Metacharacters

I think we could spend this entire course on regular expressions; they're one of my favorite subjects because I find they come in handy A LOT. It's a subject that transcends R, so you can apply this portion to other languages as well. Regular expressions are really valuable because they enable you to clean up and work with messy data, extracting what you want and leaving behind what you don't want. 

I'll begin with a more in-depth look at useful metacharacters in regular expressions, then I'll give you an exercise to test out some of those concepts. In case the terminology is confusing, the "metacharacters" in regex refer to symbols that aren't interpreted literally. For instance, the "." character isn't actually interpreted as a dot; rather, in regex it means "any character". So the following produces a match:

```{r}
my.string <- "Hi there!"
grepl(".",my.string)
```

Basically, all I did was tell R to look for any character. It'll even work on a string composed of white spaces:

```{r}
my.string <- "  "
grepl(".",my.string)
```

Some metacharacters tell R what to look for, others tell R how many times to look for it, others tell R where to look for it, and still others allow you to group or capture expressions. So, using those 4 categories as a guide, here's some common metacharacters that prove exceedingly useful:

### Character types

* `.` matches anything; it's more useful than it sounds
* `\\d` matches a digit, so any numeric character
* `\\s` matches a space
* `\\n` matches a new line character
* `\\w` matches a word, so any alphabetical character

It's also worth noting that for the `\\x` notation, a capitalized version of the same letter will match the opposite (i.e. `\\X` matches anything that isn't matched by `\\x`). Here are a couple examples to illustrate:

```{r}
my.string <- "Thishasnospaces"
grepl("\\s",my.string)
```

```{r}
grepl("\\S",my.string)
```

In the first example, we looked for a space, but found none; in the second example, we looked for anything that isn't a space and found plenty. This also brings an important concept into the foreground:

**Regular expressions are matching on the WHOLE STRING, not just a part of it**

So in the second example here, I got one `TRUE` returned for one string. There are actually 15 characters in the string, none of which are spaces, so we match the expression 15 times. But because we only need 1 match, we only get 1 `TRUE` back. This is illustrated by using `grep` and returning the matched value(s):

```{r}
grep("\\S", my.string, value = TRUE)
```

As you can see, the value we matched is the whole string, not each individual character. R is designed to run the `grep` type commands on vectors, so it's really good for finding matches in a lot of strings. If you wanted to do something with the place where you actually match, you'd have to actually use `sub` or `gsub`:

```{r}
gsub("\\S", "0",my.string)
```

When we tell R to substitute a `0` for every match, we see now that every character matched, just as we'd expect.

### Character repeats

It's oftentimes important to specify how many occurences of your pattern (or part of your pattern) you are looking for. Maybe you're looking for a phone number of 10 digits, or perhaps you have some complicated identifier that has 3 letters followed by 4 digits followed by 3 more letters. In these cases, you can use metacharacters to specify the number of occurences you're looking for. For each of these, you'll want to follow your character or expression with one of these symbols; I'll demonstrate once I've defined all of them:

* `?` matches 0 or 1 occurence
* `+` matches 1 or more occurence
* `*` matches 0 or more occurences
* `{m,n}` matches at least m, but not more than n occurences
* `{m,}` matches at least m occurences
* `{,n}` matches no more than n occurences
* `{m}` matches exactly m occurences

Here's a few examples to illustrate:

```{r}
my.string1 <- "My phone number is 8675309"
my.string2 <- "My address is 1274 King St."
my.strings <- c(my.string1, my.string2)

# Match only the address using an exact number of digits:
grepl("\\d{7}",my.strings)

# Match both by looking for 3-8 digits
grepl("\\d{3,8}",my.strings)

# Match only if there's at least 1 number 8
grepl("8+",my.strings)

# Match if there's 0 or 1 number 8
grepl("8?",my.strings)
```

### Character Placement

Where your match occurs can be really important; you might be looking for something at only the start of the string, or perhaps just at the end. There are 4 metacharacters that are useful here:

* `^` matches the pattern only at the start of a string
* `$` matches the pattern only at the end of a string
* `\\b` matches the pattern at either end
* `\\B` matches the pattern only if it's not at an end

Some examples:

```{r}
my.string1 <- 'Fred is a good friend'
my.string2 <- 'You could say my best friend is Fred'
my.strings <- c(my.string1, my.string2)

# Look for Fred only at the start
grepl("^Fred",my.strings)

# Look for Fred only at the end
grepl("Fred$",my.strings)

# Look for Fred at either end
grepl("\\bFred",my.strings)

# Look for Fred in the middle only
grepl("\\BFred",my.strings)
```

### Grouping and Capturing Expressions

There are 2 distinct concepts here that I've put together; the first is a case where you're looking for multiple possibilities at once. There's a number of metacharacters that help with that: 

* `[something]` looks for anything inside the brackets; you can use hyphens here, so `[0-9]` looks for any digit and is equivalent to `\\d`
* `[^something]` looks for anything that is NOT inside the brackets, so the "^" inside square brackets effectively means "not"
* `|` acts as an "or" to match alternative patterns
* `&` acts as an "and" to match multiple patterns

And some examples:

```{r}
my.string1 <- 'sandwiches'
my.string2 <- 'pizza'
my.strings <- c(my.string1, my.string2)

# Look for the letters "z" and "w"
grepl("[zw]", my.strings)

# Look for anything but letters that appear in "Pizza"
grepl("[^piza]", my.strings)
```

The other is useful for capturing expressions, generally for doing specific substitutions or for keeping only a sub-string:

* `(something)` allows you to group patterns to be more specific; it also allows you to capture the matched pattern, which is super useful for changing things with `sub` or `gsub`
* `\\X` pulls out matches captured by the parentheses, where X is a digit 1 or greater. So \\1 returns the first match, \\2 returns the second match, etc. 

Here's 1 example each of using the capturing to do specific substitutions, then to keep only sub-strings:

```{r}
my.string1 <- "Fred is a good friend"
my.string2 <- "You could say my best friend is Fred"
my.strings <- c(my.string1, my.string2)

# If you see "good" or "best", insert "very" in front
gsub("(good|best)","very \\1",my.strings)

# Specify the type of cheese and chocolate that I like
my.string <- "My favorite foods are cheese crackers and chocolate cake; some people don't like cheese and chocolate"
gsub("(cheese)(.+)(chocolate)","cheddar \\1\\2dark \\3",my.string)
```

There's a lot going on in this last example; see if you can make out all the components! That should do for an overview; let's get to an exercise.

## In-Class Exercises (Part 2)

KEGG is a common resource used for metabolic studies and as part of its resources, it contains a list of compounds and their metadata. I am interested in extracting all of the compounds from KEGG and creating a data frame where each compound record has the KEGG compound ID and a human-readable name. Here's what the top of the KEGG compounds file looks like:

```{r echo = FALSE}
readLines("../Datasets/keggcompounds.txt", n = 20)
```

You can also take a look at the file itself to better understand its formatting. It is located in the class Github Repository under `Datasets`. I'll post a solution after class, but I'm looking for a data frame with:

* Each compound ID
* Each corresponding name (first entry only)

As with last week's task, you might find yourself having to search around for answers along the way, and I'd highly recommend writing some pseudocode to guide you. 
