---
title: "Week 2 Materials"
author: "Matt Richards"
date: "7/13/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Week Overview

Week 2, the second and final week of the DataCamp "Intermediate R" course, was sort of a mishmash of things. After starting with some very general concepts, we had sort of a miscellaneous section to end the section. Here are the big things we covered:

* Functions: one of the most central concepts in all programming. Functions let you build modular pieces of code that you can then apply to lots of different scenarios. 

* lapply: the commonly used "loop" function; you use it to apply a function to a list or vector and you **ALWAYS** get a list in return. 
* sapply/vapply: variations on lapply that return other data types. The `sapply` simplifies it however it sees fit, so it can be a little less predictable than `vapply`, where you specify exactly the data type you want. 

*

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
