---
title: "Week 2 Slides"
author: "Matt Richards"
date: "7/13/2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Week 2 Agenda

* Quick Slide on Coding Approach
* Brief Git/Github Revisit
* Questions/Issues with Assigned Material
* Week 2 Materials: `apply` Functions
* Week 2 Materials: Regular Expressions

## Scripting Tip: Pseudocode!

My recommended problem-solving approach:

1. Make sure you completely understand what is being asked
2. Be certain of what your input looks like and what your output should look like
3. Embrace pseudocoding; write out your steps in plain English before starting to program
4. Program in small chunks, checking that you have the correct output at each step along the way
5. Be patient and don't be afraid to ask questions of your instructor or Google. 

## Pseudocoding Example

Downloading and adding a column to the "Breast Cancer" data from Week 1:

```{r echo = TRUE}
# Download the file from the URL

# Read it into a data frame and change the column names

# Create a new vector for the new column

# Cycle through each entry and:

    # Determine its identity

    # Assign a value to the new column accordingly

# Add the new column to the data frame and return it
```