---
title: "Week 1 Materials"
author: "Matt Richards"
date: "7/6/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Week Overview

This was your first taste of DataCamp and, for some of you, your introduction to control flow. When I talk about "control flow", I'm specifically talking about conditional statements and loops that allow you to dictate decision-making and repetition in your programming. In particular this week, we covered:

* Conditionals: A family of expressions that, when evaluated, result in a logical output. While typically mathematical, they don't always have to me. A non-comprehensive list of some common expressions:
    * `a > b` a greater than b
    * `a < b` a less than b
    * `a >= b` a greater than or equal to b
    * `a <= b` a less than or equal to b
    * `a != b` a not equal to b
    * `a == b` a is equal to b
    * `a %in% b` a is a member of b
    * `is.na(a)` a is NA
* `if/else/else if`: a common control structure that uses conditional to determine whether or not to execute further statements
* `for` loops: the iterative loop structure; evaluates a set of statements for every item in a specified list
* `while` loops: a conditional loop structure; it continues evaluating its set of statements until a specified conditional is no longer true
* `break` statements: allow you to terminate a loop or function, which can be especially useful when you're using variable input types

## Additional Material

### `Switch:Case`

Another useful control structure is the `switch:case` struture, which is actually quite similar to the `if/else/else if` control structure. Rather than evaluating a conditional or series of conditionals, the `switch:case` structure has a pre-defined set of cases it looks for and returns different things based on the cases it encounters. It works something like this:

```{r}
fruit <- c("apple", "banana")
color <- switch(fruit[1], 
                "apple" = "red",
                "banana" = "yellow",
                "durian" = "brown",
                "orange" = "orange")
color
```

Why use this instead of `if/else/else if`? I think it often comes down to personal preference, but here's a couple reasons:

1. It's often more concise and thus, more clear
2. It's a bit faster computationally 

On the other hand, `switch:case` tends to be less flexible than its counterpart, so it's not generally quite as widely applicable. It's designed for cases where the input (the first argument) can change and there's a whole bunch of cases you want to account for. 

### `ifelse()`

This is just a compact version of the `if/else/else if` control structure and it's quite useful for places where you just need a short test rather than drawn-out statements. Here is an example:

```{r}
x <- 1:3
ifelse(x >= 2, "Big","Little")
```

## Practice Exercises

For these exercises, you'll want to download the Breast Cancer Dataset from UCI's machine learning repository. You can go straight to their repository [here](https://archive.ics.uci.edu/ml/datasets/Breast+Cancer+Wisconsin+(Original) or, alternatively, it's on the class Github under `Datasets`. You should get the `breast-cancer-wisconsin.data` file from their data folder.

1. Read the .csv file into R so you get a data frame, then give actual titles to the variables (columns) so you know what each column is. On the UCI website, there's more information on the names of each column. 

2. The final column of the dataset specifies tumor malignancy, but does so with an odd system of 2/4 being assigned to Benign/Malignant, respectively. Using the DataCamp control structures, add a new column to the data frame that explicitly says "Benign" or "Malignant" based on the existing column. The new column should be a *factor* variable. Check your results by using `summary()` to look at your new data frame

3. Repeat exercise #2, but this time use `ifelse()` to accomplish your goal

4. Here's one from my own research: recently, we've been working with some DNA footprinting data on different chromosomes. We'd like to designate some chromosomes as something of a "training" set to establish a background level, but we want to do so randomly. We want a script that does the following:

* Creates a random integer from 1-24 (22 numbered chromosomes plus X and Y)
* Uses the random integer to designate the appropriate chromosome and pastes it to the abbreviation `chr`
* Outputs the random chromosome string (e.g. `chr8` or `chrY`)

You may find the `sample()` and `paste0()` functions useful here. 
