---
title: "Intermediate R Introduction"
author: "Matthew Richards"
date: "April 20, 2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## A Quick "About Me"

- Name
- Current Role
- Prior Programming Experience
- Experience with R
- Something Fun?

## Class Outline

- Manipulating Data with dplyr
- Control Structures
- The 'apply' Family
- R Utilities and Efficiency
- Using R Markdown for Reproducible Documents
- Git and Github for Version Control 

## Course Mechanics

As with any coding, you will learn MUCH more by doing than by watching

```{r echo=FALSE}
slices <- c(5, 20, 100)
lbls <- c("In-Class Time","DataCamp Lessons","Self-Paced Practice/Projects")
pie(slices, labels = lbls, main="Potential for Learning")

```


## Course Goals

By the end of this course, you should be able to:

- Confidently use dplyr to extract information from data frames
- Perform non-trivial operations through use of loops and conditionals
- Understand the efficiency advantages of using "apply" functions and parallel processing
- Create reproducible markdown documents with embedded R code
- Manage your code using Github repositories

## Getting help

A good rule of thumb is to proceed down this list

- Use the *"?"* or *"??"* operators
- Ask the internet
 - Google
 - StackOverflow
 - R-specific sites
- Ask your instructor
- Ask a friend who has expertise
- Use a book

## A Few Other Resources

R is learned by doing; here are some possible avenues for more "doing":

- Other DataCamp courses
- *swirl()*
- Various "Intro to R" or "R Tutorial" sites
- Start a project!

## Today's Agenda

- Intro Slides
- Creation of class Github Repo
- Review of Week 1 Assignment
- Intro to *swirl()*
- Looking into Week 2 Assignment

